---
title: "Untitled"
author: "Sherifat A-Shitu"
date: "1 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(geojsonio)
 library(sp)
library(tidyr)
library(tidyverse)
library(plotly)
```

#Overview

In 2018, Kaggle conducted an industry-wide survey that presents a truly comprehensive view of the state of data science and machine learning. The survey was live for one week in October, and after cleaning the data we finished with 23,859 responses,

This survey received 23,859 usable respondents from 147 countries and territories. If a country or territory received less than 50 respondents, we grouped them into a group named "Other" for anonymity.

```{r getdata, cache=TRUE}
##get Kaggle data in a tibble
ks_data<-tbl_df(read.csv("data/multipleChoiceResponses.csv"))

summary(ks_data)
head(ks_data)

rel_ks_data<-select(ks_data,Q1,Q3,Q12_MULTIPLE_CHOICE,Q13_Part_1:Q13_Part_15,Q14_Part_1:Q14_Part_11,Q15_Part_1:Q15_Part_7,Q16_Part_1:Q16_Part_18,Q17,Q18,Q19_Part_1:Q19_Part_19,Q20,Q21_Part_1:Q21_Part_13,Q22, Q27_Part_1:Q27_Part_20, Q28_Part_1:Q28_Part_43, Q29_Part_1:Q29_Part_28, Q30_Part_1:Q30_Part_25,Q31_Part_1:Q31_Part_12, Q37)

#set column names to first row values
colnames(rel_ks_data) <- rel_ks_data%>%
                            slice( 1) %>%
                              unlist(., use.names=FALSE)

nrow( filter(rel_ks_data, Q13_Part_1=="Jupyter/IPython"))

rel_ks_data %>% 
  rename("Gender" = !!names(.[1]),
            "Country" = !!names(.[2]))%>%
  rowid_to_column("ID")%>%
    gather(key=question,value = response,-ID,-Gender,-Country)%>%
  dcast( ID+Gender+Country ~ question)

#get and clean country data
country_data<-data.frame(table(ks_data$Q3))
colnames(country_data)<-c("Country","Num Of Response")
country_data$Country<-as.character(country_data$Country)
country_data$Country[59]<-"Vietnam"
country_data$Country[57]<-"United Kingdom"
country_data$Country[44]<-"North Korea"
country_data$Country[25]<-"Iran"
country_data$Country[25]<-"Hong Kong S.A.R."
```

```{r reports,cache=TRUE}

#num of countrires
nlevels(ks_data$Q3)
levels(ks_data$Q3)

#Age distribution
age<-ks_data%>%count(Q2, sort = TRUE)%>%mutate(freq = (n / sum(n))*100)
round(age[1,]$freq,0)

#num responses
nrow(ks_data)-1

#gender dist
gender<-ks_data%>%count(Q1, sort = TRUE)%>%mutate(freq = (n / sum(n))*100)

#top 5 countries
t5c<-ks_data%>%
  count(Q3, sort = TRUE)%>%
  mutate(freq = (n / sum(n))*100)%>%
  filter(Q3!="Other")%>%
  slice(1:5)%>%arrange(n)

#no axes
ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)
#pie chart
plot_ly(t5c, values= c(t5c$n), labels = ~reorder(c(as.character(t5c$Q3)),c(t5c$n)),
        type = 'pie',textposition = 'inside',
        textinfo = 'label+percent+value',
        insidetextfont = list(color = 'purple'),
        hoverinfo = 'text',
        text = ~format(t5c$n, big.mark = ","),
        marker = list(colors = blues9 ,
                      line = list(color = 'lightblue', width = 1)),
        showlegend = FALSE) %>%
  layout(width=350, height=350,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

#bar chart
plot_ly(t5c, x = c(t5c$n), y = reorder(c(as.character(t5c$Q3)),c(t5c$n)),
            type = 'bar', orientation = 'h' , 
            marker = list(color = 'rgba(38, 24, 74, 0.8)',
                          line = list(color = 'rgba(122, 120, 168, 1.0)', width = 1))) %>%
      layout(yaxis = list(showgrid = FALSE, showline = FALSE, showticklabels = TRUE, domain= c(0, 0.85)),
             xaxis = list(zeroline = FALSE, showline = FALSE, showticklabels = TRUE, showgrid = TRUE)) %>%
      add_annotations(xref = 'x1', yref = 'y',
                      x = t5c$n /2 ,  y = reorder(c(as.character(t5c$Q3)),c(t5c$n)),
                      text = paste("<b>",as.character(t5c$Q3),"\n",formatC(t5c$n, big.mark = ","),"</b>"),
                      font = list(family = 'Arial', size = 10, color="white" ),
                      showarrow = FALSE)%>%
      layout(barmode = 'stack',width = 250, height = 350,
             xaxis = ax,
             yaxis = ax)


ggplot(t5c, aes(x=Q3, y=n)) +
  geom_bar(stat='identity') +
  coord_flip()

c(t5c$n)


##programming language usage

selected=c("Nigeria","Belgium")
pl_usage<- ks_data%>%
            filter(Q3!= c("I do not wish to disclose my location" ,"In which country do you currently reside?"),
                   Q3==selected )%>%
            select(Q16_Part_1:Q16_Part_18)%>%
            gather(x, value, Q16_Part_1:Q16_Part_18)%>%
            group_by(x)%>%
            tally(value !="")%>%
              mutate(id=gsub("Q16_Part_", "", x))
            

summary(pl_usage)



plang<-c("1"= "Python",
			   "2"="R",
			   "3"="SQL",
			   "4"="Bash",
			   "5"= "Java",
			   "6"="JavaScript/Typescript",
			   "7"="Visual Basic/VBA",
			   "8"="C/C++",
			   "9"="MATLAB",
			   "10"="Scala",
         "11"="Julia",
         "12"="Go",
         "13"="C#/.NET",
         "14"="PHP",
         "15"="Ruby",
         "16"="SAS/STATA",
			   "17"="None",
			   "18"="Other"
)
pl_usage$id<-recode_factor(pl_usage$id, !!!plang)
pl_usage$id<-as.character(pl_usage$id)
pl_usage$id[15]="Javascript/Typescript"
pl_usage$id<-as.factor(pl_usage$id)

ks_data%>%
  filter(Q3!= c("I do not wish to disclose my location" ,"In which country do you currently reside?"),
         Q17!=c("","What specific programming language do you use most often? - Selected Choice"))%>%
  group_by(Q17)%>%
  summarize(count = n())%>% rename(id=Q17)%>%
  slice(2:18)



plang_data <-full_join(pl_usage, ks_data%>%
          filter(Q3!= c("I do not wish to disclose my location" ,"In which country do you currently reside?"),
                    Q3==selected,
                 Q17!=c("","What specific programming language do you use most often? - Selected Choice"))%>%
          group_by(Q17)%>%
          summarize(count = n())%>%dplyr::rename(id=Q17)%>%
          slice(2:18),by=tolower("id"))%>%replace_na(list(count = 0))


full_join(pl_usage, ks_data%>%
            filter(Q3!= c("I do not wish to disclose my location" ,"In which country do you currently reside?"),
                            Q3==selected,
                  Q17!=c("","What specific programming language do you use most often? - Selected Choice"))%>%
            group_by(Q17)%>%
            summarize(count = n())%>% rename(id=Q17)%>%
            slice(2:18),by=tolower("id"))%>%replace_na(list(count = 0))

p1 <- plot_ly(plang_data,x = ~n, y = ~reorder(id, n), 
              name = 'What programming languages do you use on a regular basis?',
              type = 'bar', orientation = 'h',
              marker = list(color = 'white',
                            line = list(color = 'rgba(50, 171, 96, 1.0)', width = 1))) %>%
  layout(yaxis = list(showgrid = FALSE, showline = FALSE, showticklabels = TRUE, domain= c(0, 0.85)),
         xaxis = list(zeroline = FALSE, showline = FALSE, showticklabels = TRUE, showgrid = TRUE)) %>%
  add_annotations(xref = 'x1', yref = 'y',
                  x = ~n/3,  y = ~id,
                  text = ~formatC(plang_data$n, big.mark ="," ),
                  font = list(family = 'Arial', size = 12, color = 'rgb(50, 171, 96)'),
                  showarrow = FALSE)

p2 <- plot_ly(plang_data,x = ~count, y = ~reorder(id, n), 
              name = 'What specific programming language do you use most often?',
              type = 'bar', orientation = 'h',
              marker = list(color = 'white',
                            line = list(color = 'rgba(128, 0, 128, 1.0)', width = 1))) %>%
 layout(yaxis = list(showgrid = FALSE, showline = FALSE, showticklabels = FALSE, domain= c(0, 0.85)),
         xaxis = list(zeroline = FALSE, showline = FALSE, showticklabels = TRUE, showgrid = TRUE)) %>%
  add_annotations(xref = 'x1', yref = 'y',
                  x = ~count/3,  y = ~id,
                  text = ~formatC(plang_data$count, big.mark ="," ),
                  font = list(family = 'Arial', size = 12, color = 'rgb(128,0,128)'),
                  showarrow = FALSE)


subplot(p1, p2) %>%
  layout(title = 'Most popular ML/DS Programming Language',
         legend = list(x = 0.029, y = 1.038,
                       font = list(size = 10)),
         margin = list(l = 100, r = 20, t = 70, b = 70),
         paper_bgcolor = 'rgb(248, 248, 255)',
         plot_bgcolor = 'rgb(248, 248, 255)') %>%
  add_annotations(xref = 'paper', yref = 'paper',
                  x = -0.14, y = -0.15,
                  text = paste('Data Source: https://www.kaggle.com/kaggle/kaggle-survey-2018#multipleChoiceResponses.csv'),
                  font = list(family = 'Arial', size = 10, color = 'rgb(150,150,150)'),
                  showarrow = FALSE)



```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(maps);library(leaflet);library(rgdal)
#use geojsonio or rgdal (packages) to read GeoJSON country data as sp objects.
#data source: https://raw.githubusercontent.com/datasets/geo-countries/master/data/countries.geojson

countries <- geojsonio::geojson_read("./data/countries.geojson",  what = "sp")
#add num_of_responses per country from country_data to 'countries' sp object
countries@data = data.frame(countries@data, country_data[match(countries@data[,"ADMIN"], country_data[,"Country"]),])

#remove NA Values
countries_complete<-sp.na.omit(countries, col.name = "Num.Of.Response") 
countries@data <-countries@data[complete.cases(countries@data),] 

##create col pallete
pal<- colorNumeric("Blues", country_data$`Num Of Response`, n = 4)


leaflet(countries_complete) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(Num.Of.Response),
    label = ~paste0(Country, ": ", formatC(Num.Of.Response, big.mark = ","))) %>%
  addLegend(pal = pal, values = ~Num.Of.Response, opacity = 1.0 )


leaflet() %>% setView(lng = -98.583, lat = 39.833, zoom = 2) %>%
  addTiles() %>%
  addTopoJSON(topoData, weight = 1, color = "#444444", fillColor = pal(country_data$`Num Of Response`))

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
